Pulling oral and IV data from CvTdb

### 2025-04-11 data-raw/pulling_oral_iv_data_cvtdb.Rmd

@@

quick_message(text = c("Filtering out:",
                       "Missing {.code analyte_dtxsid}",
                       "Radiolabeling experiments",
                       "Data where the analyzed chemicals and dosed chemicals are different"))
```

@@

cvt_df <- cvt_df %>%
  filter(!(is.na(conc) & is.na(conc_original))) %>%
  group_by(fk_study_id) %>%
  filter(n() > 1) %>%
  ungroup()
###
```

@@

# Checking other_study_identifier == S0916
quick_message(text = c(
  "Data Modification:",
  paste("For study 1713 and 1714, tamoxifen was administered at 5mL/kg, ",
        "meaning these doses are 5/1000th of current value to convert ug/mL -> mg/kg.")
))
# In CEBS "S0916" the Dose is 300,100 ug/mL with 5mL/kg administered by gavage (the only data left)
# Making it 5/1000 mg/kg x either doses in study_ids 1714,1713 (the latter was filtered out prior to this)
cvt_df <- cvt_df %>%
  mutate(
    invivPK_dose_level = if_else(
      fk_study_id %in% c("1714", "1713"),
      5/1000 * dose_level_normalized,
      dose_level_normalized
    )
  )

@@

quick_message("CEBS datasets S0976 and S0328 contain data on LOQ or ELOQ")
# For study S0976, Caffeine, L-ephedrine, and psuedoephedrine 
quick_message(c(
  "Data Modification:",
  "Adjusting LOQ values for L-ephedrine to 3.39 as stated in CEBS study S0976.",
  "Set LOQ units to ng/mL as detailed in the same CEBS study"
))
# Here I create new columns for loq and loq_units b/c I have modified them.
cvt_df <- cvt_df %>%
  mutate(invivPK_loq = if_else(
    other_study_identifier %in% "S0976",
    case_when(
      analyte_name_original %in% "caffeine" ~ 4.00,
      analyte_name_original %in% "l-ephedrine" ~ 3.39,
      analyte_name_original %in% "pseudoephedrine" ~ 4.43
    ),
    loq),
    invivPK_loq_units = if_else(
      other_study_identifier %in% "S0976",
      "ng/mL",
      loq_units
    ))

# Some pentachlorophenol studies
quick_message(c(
  "Data Modification:",
  "Looked at ELOQ in CEBS study S0328 and found it to be around 1",
  "This only is relevant for {.code study_id} 1779"
))

cvt_df <- cvt_df %>%
  mutate(
  invivPK_loq = if_else(
    fk_study_id %in% "1779",
    1, invivPK_loq),
  invivPK_loq_units = if_else(
    other_study_identifier %in% "S0328",
    "ug/mL",
    invivPK_loq_units
  ))

# quick_stats(cvt_df)

@@

# quick_stats(cvt_df)
quick_message("Uniqe LOQ units:")
table(cvt_df$invivPK_loq_units)

@@

# This is the final filtering so that the dataset is most consistent with the
# most current near future pull in the new year
cvt_df <- cvt_df %>%
  filter(invivPK_conc > 0)

@@

glimpse(cvt_df, width = 80)
quick_stats(cvt_df)
```

@@

```{r diff_log_function}
# Now I want to write a quick function that will log relevant changes
# from previous 'cvt' objects
devtools::load_all() # So this loads the previous cvt object
diff_log_cvt <- function(new_cvt,
                         key_id = "conc_time_id",
                         grp_ids = c("analyte_dtxsid",
                                     "species")) {
  stopifnot(
    key_id %in% colnames(cvt),
    key_id %in% colnames(new_cvt),
    all(grp_ids %in% colnames(cvt)),
    all(grp_ids %in% colnames(new_cvt))
  )
  if (length(unique(new_cvt[[key_id]])) != NROW(new_cvt)) {
    stop("key_id is not not a column that uniquely identifies this data.")
  }

    ### -----Checking columns
    diff_col_add <- setdiff(names(new_cvt), names(cvt)) # Columns added in new_cvt
    diff_col_sub <- setdiff(names(cvt), names(new_cvt)) # Columns removed in new_cvt
    diff_col_common <- intersect(names(cvt), names(new_cvt))

    # Left off here TODO
    if (length(c(diff_col_add, diff_col_sub)) == 0) {
      msg_to_write <- c(msg_to_write,
                        "No changes in column names.")
    } else if (length(diff_col_add) > 0 && length(diff_col_sub) > 0) {
      # Check renaming
      common_row_cvt <- cvt[diff_row_common, ]
      common_row_new <- new_cvt[diff_row_common, ]

    if (!identical(common_row_cvt[diff_col_common],
                   common_row_new[diff_col_common])) {
      msg_to_write <- c(msg_to_write,
                        "Data for same observation key_ids have changed.")
    }
    msg_to_write <- c(msg_to_write,
                      "================================\n")
  }
  return(cat(msg_to_write, sep = "\n"))
}
```

@@

```{r saving_data, include=FALSE, eval=FALSE}
cvt_date <- Sys.Date()
cvt <- cvt_df
save(cvt, file = "data/cvt.rda", compress = "bzip2")
save(cvt_date, file = "data/cvt_date.rda")
```

>>>>

```{r diff_log_function}
# Now I want to write a quick function that will log relevant changes
# from previous 'cvt' objects
devtools::load_all() # So this loads the previous cvt object
diff_log_cvt <- function(new_cvt,
                         key_id = "conc_time_id",
                         grp_ids = c("analyte_dtxsid",
                                     "species"),
                         notes = character(0L)) {
  stopifnot(
    key_id %in% colnames(cvt),
    key_id %in% colnames(new_cvt),
    all(grp_ids %in% colnames(cvt)),
    all(grp_ids %in% colnames(new_cvt))
  )
  if (length(unique(new_cvt[[key_id]])) != NROW(new_cvt)) {
    stop("key_id is not not a column that uniquely identifies this data.")
  }

    ### -----Checking columns
    diff_col_add <- setdiff(names(new_cvt), names(cvt)) # Columns added in new_cvt
    diff_col_sub <- setdiff(names(cvt), names(new_cvt)) # Columns removed in new_cvt
    diff_col_common <- intersect(names(cvt), names(new_cvt))
    common_row_cvt <- cvt[cvt[[key_id]] %in% diff_row_common, ]

    common_row_cvt <- common_row_cvt[order(common_row_cvt[[key_id]]), ]

    common_row_new <- new_cvt[new_cvt[[key_id]] %in% diff_row_common, ]
    common_row_new <- common_row_new[order(common_row_new[[key_id]]), ]

    if (length(c(diff_col_add, diff_col_sub)) == 0) {
      msg_to_write <- c(msg_to_write,
                        "No changes in column names.")
    } else if (length(diff_col_add) > 0 && length(diff_col_sub) > 0) {
      # Check renaming

    if ((NROW(
      setdiff(
        common_row_cvt[diff_col_common], common_row_new[diff_col_common]
      )) > 0
    ) ||  (NROW(
      setdiff(
        common_row_new[diff_col_common], common_row_cvt[diff_col_common]
      )) > 0
    )) {
      msg_to_write <- c(msg_to_write,
                        "Data for same observation key_ids have changed.")

      # Which columns have different data?
      test_cols <- setdiff(diff_col_common, key_id)
      for (i in test_cols) {
        this_diff <- setdiff(common_row_new[c(key_id, i)],
                             common_row_cvt[c(key_id, i)])
        that_diff <- setdiff(common_row_cvt[c(key_id, i)],
                             common_row_new[c(key_id, i)])

        if (NROW(this_diff) > 0) {
          msg_to_write <- c(
            msg_to_write,
            paste0(c(i, "had", NROW(this_diff), "differences."), collapse = " ")
          )

          this_grp <- dplyr::distinct(common_row_new[
            common_row_new[[key_id]] %in% this_diff[[key_id]],
            grp_ids
          ])
          this_grp <- paste0(this_grp[1,1:2], collapse = " ")

          # Print the differences (if below a certain number of rows)
          msg_to_write <- c(
            msg_to_write,
            paste0(
              "Example:",
              paste(this_grp, collapse = " "),
              collapse = " "
            ),
            paste0(c(i, "\t new:", this_diff[1,2], "<--\t old:", that_diff[1,2]),
                   collapse = " "), "\n"
          )
        }
      }
    }
    msg_to_write <- c(msg_to_write,
                      notes,
                      "================================\n")
  }
  return(cat(msg_to_write, sep = "\n"))
}

@@

quick_message(text = c("Filtering out:",
                       "Missing {.code analyte_dtxsid}",
                       "Radiolabeling experiments",
                       "Data where the analyzed chemicals and dosed chemicals are different"))
diff_log_cvt(cvt_df)
```

@@

# Filtering out entries with no concentration information available
cvt_df <- cvt_df %>%
  filter(!(is.na(conc) & is.na(conc_original))) %>%
  group_by(fk_study_id) %>%
  filter(n() > 1) %>%
  ungroup()
###
```

@@

cvt_df <- cvt_df %>%
  mutate(invivPK_dose_level = dose_level_normalized,
         invivPK_loq = loq,
         invivPK_loq_units = loq_units,
         invivPK_conc_sd = conc_sd,
         invivPK_conc = conc)
diff_log_cvt(cvt_df)

# Checking other_study_identifier == S0916
quick_message(text = c(
  "Data Modification:",
  paste("For study 1713 and 1714, tamoxifen was administered at 5mL/kg, ",
        "meaning these doses are 5/1000th of current value to convert ug/mL -> mg/kg."
),
 "This entry has since been removed from CvT as of April 2025"
)
)

# In CEBS "S0916" the Dose is 300,100 ug/mL with 5mL/kg administered by gavage (the only data left)
# Making it 5/1000 mg/kg x either doses in study_ids 1714,1713 (the latter was filtered out prior to this)
# cvt_df <- cvt_df %>%
#   mutate(
#     invivPK_dose_level = if_else(
#       fk_study_id %in% c("1714", "1713"),
#       5/1000 * dose_level_normalized,
#       dose_level_normalized
#     )
#   )

@@

quick_message("CEBS datasets S0976 and S0328 contain data on LOQ or ELOQ")
# For study S0976, Caffeine, L-ephedrine, and psuedoephedrine 
quick_message(c(
  "Data Modification:",
  "Adjusting LOQ values for L-ephedrine to 3.39 as stated in CEBS study S0976.",
  "Set LOQ units to ng/mL as detailed in the same CEBS study.",
  "Caffeine and Pseudoephedine have been removed as of April 2025."
))
# Here I create new columns for loq and loq_units b/c I have modified them.
cvt_df <- cvt_df %>%
  mutate(invivPK_loq = if_else(
    other_study_identifier %in% "S0976",
    case_when(
      # analyte_name_original %in% "caffeine" ~ 4.00,
      analyte_name_original %in% "l-ephedrine" ~ 3.39,
      # analyte_name_original %in% "pseudoephedrine" ~ 4.43
    ),
    loq),
    invivPK_loq_units = if_else(
      other_study_identifier %in% "S0976",
      "ng/mL",
      loq_units
    ))

# Some pentachlorophenol studies
quick_message(c(
  "Data Modification:",
  "Looked at ELOQ in CEBS study S0328 and found it to be around 1",
  "This only is relevant for {.code study_id} 1779.",
  "This record is no longer present as of April 2025."
))

# cvt_df <- cvt_df %>%
#   mutate(
#   invivPK_loq = if_else(
#     fk_study_id %in% "1779",
#     1, invivPK_loq),
#   invivPK_loq_units = if_else(
#     other_study_identifier %in% "S0328",
#     "ug/mL",
#     invivPK_loq_units
#   ))

# quick_stats(cvt_df)
diff_log_cvt(cvt_df)
```

@@

# quick_stats(cvt_df)
quick_message("Uniqe LOQ units:")
table(cvt_df$invivPK_loq_units)
diff_log_cvt(cvt_df)

@@

# This is the final filtering so that the dataset is most consistent with the
# most current near future pull in the new year
# This will include some zero values, but should work itself out during preprocessing
cvt_df <- cvt_df %>%
  filter(invivPK_conc >= 0)

@@

glimpse(cvt_df, width = 80)
quick_stats(cvt_df)
diff_log_cvt(cvt_df, 
             notes = "Relaxed conc > 0 filter to conc >= 0.")

@@

@@

```{r saving_data, include=FALSE, eval=FALSE}
cvt_date <- Sys.Date()
cat(diff_log_cvt(cvt_df,
                 notes = "Relaxed conc > 0 filter to conc >= 0."),
    file = "data-raw/cvt_changelog.txt",
    append = TRUE)

cvt <- cvt_df
save(cvt, file = "data/cvt.rda", compress = "bzip2")
save(cvt_date, file = "data/cvt_date.rda")
